Co-methylation

### 2015-04-14 Rmarkdown/comethylation.Rmd
Problems with simple log odds-ratios.

\FloatBarrier
\subsubsection{$NIL \geq 0$}\label{sec:cometh_nil_geq0}

@@

>>>>

\FloatBarrier

\subsubsection{$NIL \geq 0$}\label{sec:cometh_nil_geq0}

@@

# Problem with LOR estimates

The simple estimator of within-fragment co-methylation is biased when the marginal probability of methylation is close to the boundary (0 or 1) unless the 2-tuple is sequenced at very high depth. I demonstrate this by simulating methylation patterns at 2-tuples as independent $binomail(1, p)$ random variables and showing that the simple LOR estimator is positively biased.

```{r}
# Function to simulate n 2-tuples with d sequencing coverage and marginal 
# probabilities p1, p2
f <- function(n, d, p1, p2) {
  sapply(seq_len(n), function(i, d, p1, p2) {
  z1 <- factor(rbinom(d, 1, p1), levels = c(0, 1))
  z2 <- factor(rbinom(d, 1, p2), levels = c(0, 1))
  x <- xtabs(~ z2 + z1, drop.unused.levels = FALSE)
  log2(x[1, 1] + 0.5) + log2(x[2, 2] + 0.5) - 
    log2(x[1, 2] + 0.5) - log2(x[2, 1] + 0.5)
  }, d = d, p1 = p1, p2 = p2)
}

# Vary p1 with p2 = p1 at 30x coverage
d <- 30
n <- 1000
p <- c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
x <- lapply(p, function(p) {
  f(n, d, p, p)
})
par(mfrow = c(3, 3))
mapply(function(x, p) {
  hist(x, main = paste0("p1 = p2 = ", p), xlim = c(-10, 10), breaks = 20)
  abline(v = 0, col = "red", lty = 2)
  }, x = x, p = p)
title(paste("d = ", d), outer = TRUE, line = -1)

# Vary p1 with p2 = 1 - p1
d <- 30
n <- 1000
p <- c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
x <- lapply(p, function(p) {
  f(n, d, p, 1 - p)
})
par(mfrow = c(3, 3))
mapply(function(x, p) {
  hist(x, main = paste0("p1 = ", p, ", p2 = ", 1 - p), xlim = c(-10, 10), breaks = 20)
  abline(v = 0, col = "red", lty = 2)
  }, x = x, p = p)
title(paste("d = ", d), outer = TRUE, line = -1)

# Vary p1 with p2 = 0.5
d <- 30
n <- 1000
p <- c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
x <- lapply(p, function(p) {
  f(n, d, p, 0.5)
})
par(mfrow = c(3, 3))
mapply(function(x, p) {
  hist(x, main = paste0("p1 = ", p, ", p2 = ", 0.5), xlim = c(-10, 10), breaks = 20)
  abline(v = 0, col = "red", lty = 2)
  }, x = x, p = p)
title(paste("d = ", d), outer = TRUE, line = -1)

# Vary p1 with p2 = p1 at 1000x coverage
d <- 1000
n <- 1000
p <- c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
x <- lapply(p, function(p) {
  f(n, d, p, p)
})
par(mfrow = c(3, 3))
mapply(function(x, p) {
  hist(x, main = paste0("p1 = p2 = ", p), xlim = c(-10, 10), breaks = 20)
  abline(v = 0, col = "red", lty = 2)
  }, x = x, p = p)
title(paste("d = ", d), outer = TRUE, line = -1)

```
